---
title: "Homework 8"
author: "Melissa Yee"
date: "2025-04-11"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(mosaic)
```

Github Link: https://github.com/myee240/HW8 

# **Problem 1: Regression warm up**

```{r echo=FALSE, message=FALSE, results = 'hide'}
creatinine <- read.csv("creatinine.csv")
lm_creatinine <- lm(creatclear ~ age, data = creatinine)
coef(lm_creatinine) %>% round(2)
```
Fitted equation: 147.81  + (-0.62)(55) = 113.71 mL/min 

**Part A**: For a 55-year old, we would expect a creatinine clearance rate of 113.71 mL/min. This was found by plugging 5 into the regression equation: 147.81 - 0.62 * age 

**Part B**: The slope β1 is -0.62, meaning the clearance decreases by 0.62 mL/min for each additional year of age. 

**Part C**: 

Age 40 Prediction: 147.81 - 0.62 * 40 = 123.01

Age 40 Actual = 135 

Age 40 Residual = 135 - 123.01 = +11.99 


Age 60 Prediction: 147.81 - 0.62 * 60 = 110.61

Age 60 Actual = 112 

Age 60 Residual = 112 - 110.61 = +1.39

The 40-year-old is healthier for their age, with a clearance rate 11.99 mL/min above expected, compare to only 1.39 mL/min above expected for the 60-year-old. I determined this by calculating the residuals and comparing each person's actual rate to the predicted rate from the regression model. 


# **Problem 2: Modeling disease growth**

```{r echo=FALSE, message=FALSE, results = 'hide'}
covid <- read.csv("covid.csv")
covid <- subset(covid, deaths > 0)

italy <- subset(covid, country == "Italy")
spain <- subset(covid, country == "Spain")

lm_italy <- lm(log(deaths) ~ days_since_first_death, data=italy)
boot_lm_italy = do(1000)*lm(log(deaths) ~ days_since_first_death, data=resample(italy))
head(boot_lm_italy)

boot_lm_spain <- do(1000)*lm(log(deaths) ~ days_since_first_death, data=resample(spain))
lm_spain <- lm(log(deaths) ~ days_since_first_death, data=spain)
boot_lm_spain = do(1000)*lm(log(deaths) ~ days_since_first_death, data=resample(spain))

# Italy growth rate CI 
italy_growth_rate_ci <- confint(boot_lm_italy, level=0.95)
head(italy_growth_rate_ci)

# Spain growth rate CI 
spain_growth_rate_ci <- confint(boot_lm_spain, level=0.95)
head(spain_growth_rate_ci)


# Italy doubling time CI 
doubling_italy <- 70 / (boot_lm_italy$days_since_first_death * 100)
quantile(doubling_italy, c(0.025, 0.975))

# Spain doubling time CI
doubling_spain <- 70 / (boot_lm_spain$days_since_first_death * 100)
quantile(doubling_spain, c(0.025, 0.975))

```
**Part A**: 

An estimated growth rate for Italy: 0.183 

Confidence Interval: (0.160, 0.208)


An estimated doubling time for Italy: 3.8

Confidence Interval: (3.4, 4.4) days


I found the estimated death growth rate for Italy to be 0.183, or 18.3% and computed a 95% confidence interval of (0.160, 0.208). 
Since the task was to fit an exponential growth model, I fitted our model using a logarithmic scale for the y variable. This linear model is fit for Italy using log-transformed death counts as the response variable and days since the first death as the predictor. Next, I performed 1000 bootstrap samples of the data to compute 95% confidence intervals for the growth rates. Then, I found the estimated doubling time for Italy to be 3.8 days and computed a 95% confidence interval of (3.4, 4.4). I calculated the doubling time of the daily death total Italy using the "Rule of 70" and additionally bootstrapped this estimate as well to create another 95% confidence interval. 

**Part B**: 

An estimated growth rate for Spain: 0.276

Confidence Interval: (0.238, 0.320)


An estimated doubling time for Spain: 2.5

Confidence Interval: (2.2, 2.9) days


Same procedure and analysis from Part B on Italy was performed on Spain. 

**Part C**
``` {r echo=FALSE, message=FALSE, results = 'hide', warning=FALSE}
# Basic line plot
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Daily Reported COVID-19 Deaths Over Time",
    x = "Days Since First Reported Death",
    y = "Number of Deaths",
    color = "Country"
  ) +
  theme_minimal() 

```


This line graph shows reported daily deaths over time (using days_since_first_death) in Italy and Spain.

# **Problem 3: Price elasticity of demand**

``` {r echo=FALSE, message=FALSE, results = 'hide', include=FALSE}
milk <- read.csv("milk.csv")

ggplot(milk) +
   geom_point(aes(x=price, y=sales))

ggplot(milk) +
   geom_point(aes(x=log(price), y=log(sales)))

lm_milk = lm(log(sales) ~ log(price), data=milk)

coef(lm_milk)

confint(lm_milk)
```

My estimated elasticity is about -1.62. This means that when the price of milk increases by 1%, consumers want to buy about 1.62% less of it, on average.

An estimated elasticity of demand: -1.62
Confidence Interval: (-1.779, -1.458)

A characteristic feature of a power law is that it looks linear when plotting log(y) versus log(x). When I fit a linear model for log(sales) vs log(price), I can estimate the intercept and log(price) slope of our fitted equation. Finally, I found the confidence interval for the price elasticity of demand at 95% confidence using bootstrapped samples of the linear model. 
